---
title: "Homework 6"
author: "Mukta Patwari"
date: "2025-11-11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

library(readr)
library(broom)
library(modelr)
library(patchwork)

set.seed(1234)
```

## Problem 1

Loading the homicide data

```{r}
homicide_data =
  read_csv("homicide-data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    solved = factor(
  case_when(
    disposition %in% c("Closed without arrest", "Open/No arrest") ~ "Unsolved",
    TRUE ~ "Resolved"),
  levels = c("Unsolved", "Resolved"))
  ) %>% 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
  ) %>% 
  mutate(
    victim_age = as.numeric(victim_age)
  )
```

Odds ratios for Baltimore

```{r}
baltimore_df =
  homicide_data %>% 
  filter(city_state == "Baltimore, MD")

baltimore_fit =
  baltimore_df %>% 
  glm(
    formula = solved ~ victim_age + victim_sex + victim_race,
    data = .,
    family = binomial
  ) 

baltimore_fit %>% 
  broom::tidy(conf.int = TRUE) %>% 
  mutate(
    OR = exp(estimate), 
    OR_conf_low = exp(conf.low), 
    OR_conf_high = exp(conf.high)) %>%
  filter(term == "victim_sexMale") %>%
  select(term, OR, OR_conf_low, OR_conf_high) %>% 
  knitr::kable(digits = 3)
```

Odds ratios for all cities

```{r}
cities_fit =
  homicide_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    fit = map(data, ~ glm(
      solved ~ victim_age + victim_sex + victim_race,
      data = .x,
      family = binomial
    )),
    
    tidy = map(fit, ~ 
      broom::tidy(.x, conf.int = TRUE) %>%
        mutate(
          OR = exp(estimate),
          OR_conf_low = exp(conf.low),
          OR_conf_high = exp(conf.high)
        ) %>%
        filter(term == "victim_sexMale") %>%
        select(term, OR, OR_conf_low, OR_conf_high)
    )
  ) %>%
  unnest(tidy)
```

Plot

```{r}
cities_fit %>%
  arrange(OR) %>%
  mutate(city_state = factor(city_state, levels = city_state)) %>%
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = OR_conf_low, ymax = OR_conf_high), width = 0.2) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(
    title = "Adjusted Odds Ratios for Solved Homicides",
    x = "City",
    y = "Adjusted Odds Ratio (Male vs Female)"
  )
```

## Problem 2

```{r}
library(p8105.datasets)
data("weather_df")
```

**Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution for your estimates, and describe these in words.**

```{r}
bootstrap_weather =
  weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    model = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    r_squared = map_dbl(model, ~ glance(.x)$r.squared),
    beta_ratio = map_dbl(model, ~ {
      coefs = tidy(.x)
      b1 = coefs$estimate[coefs$term == "tmin"]
      b2 = coefs$estimate[coefs$term == "prcp"]
      b1 / b2
    })
  )

r_squared_plot =
  bootstrap_weather %>%
  ggplot(aes(x = r_squared)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Bootstrap Distribution of R-squared", 
    x = "R-squared", 
    y = "Density")

beta_ratio_plot =
  bootstrap_weather %>% 
  ggplot(aes(x = beta_ratio)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Bootstrap Distribution of Beta 1 / Beta 2",
    x = "Beta 1 / Beta 2",
    y = "Density"
  )

r_squared_plot + beta_ratio_plot
```

This plot shows

**Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval.**

```{r}
bootstrap_weather %>% 
  summarise(
    rsquared_lower = quantile(r_squared, 0.025),
    rsquared_upper = quantile(r_squared, 0.975),
    betaratio_lower = quantile(beta_ratio, 0.025),
    betaratio_upper = quantile(beta_ratio, 0.975)
  ) %>%
  pivot_longer(
    everything(),
    names_to = c("Estimates", "CI"),
    names_sep = "_"
  ) %>%
  pivot_wider(
    names_from = CI,
    values_from = value
  ) %>%
  knitr::kable(digits = 3)
```

## Problem 3

**Load and clean the data for regression analysis (i.e. use appropriate variable names, convert numeric to factor where appropriate, check for the presence of missing data, etc.)**

```{r}
birthweight_df =
  read_csv("birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    babysex = factor(
      case_match(
        babysex,
        1 ~ "male",
        2 ~ "female"
      ),
      levels = c("male", "female")
    ),
    frace = factor(
      case_match(
        frace,
        1 ~ "White",
        2 ~ "Black",
        3 ~ "Asian",
        4 ~ "Puerto Rican",
        8 ~ "Other",
        9 ~ "Unknown"
      ),
      levels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")
    ),
    malform = factor(
      case_match(
        malform,
        0 ~ "absent",
        1 ~ "present"
      ),
      levels = c("absent", "present")
    ),
    mrace = factor(
      case_match(
        mrace,
        1 ~ "White",
        2 ~ "Black",
        3 ~ "Asian",
        4 ~ "Puerto Rican",
        8 ~ "Other",
        9 ~ "Unknown"
      ),
      levels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")
    )
  )
```

**Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot**

Based on preexisting literature, some factors commonly associated with low birthweight are [smoking during pregnancy](https://link.springer.com/article/10.1186/s12884-024-06487-y), [maternal age at pregnancy](https://academic.oup.com/aje/article/186/11/1219/3858068), and [mother's race](https://www.sciencedirect.com/science/article/abs/pii/S0091743518303311?via%3Dihub). Therefore, the proposed model is that factors such as smoking, maternal age, and maternal race affect birthweight. The following plot below shows this regression model. 

```{r}
model_bwt = lm(bwt ~ smoken + momage + mrace, data = birthweight_df)

birthweight_df %>% 
  add_predictions(model_bwt) %>% 
  add_residuals(model_bwt) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.6) +
  labs(
    x = "Predicted Birthweight",
    y = "Residuals",
    title = "Fitted Values vs Residuals for Birthweight"
  )
```

**Compare your model to two others: 1) One using length at birth and gestational age as predictors (main effects only), 2) One using head circumference, length, sex, and all interactions (including the three-way interaction) between these. Make this comparison in terms of the cross-validated prediction error; use `crossv_mc` and functions in `purrr` as appropriate.**

```{r}
cv_df =
  crossv_mc(birthweight_df, n = 100)

cv_results = 
  cv_df %>%
  mutate(
    model1 = map(train, ~ lm(bwt ~ smoken + momage + mrace, data = .x)),
    model2 = map(train, ~ lm(bwt ~ blength + gaweeks, data = .x)),
    model3 = map(train, ~ lm(bwt ~ bhead * blength * babysex, data = .x)),
    model1_rmse = map2_dbl(model1, test, ~ rmse(.x, .y)),
    model2_rmse = map2_dbl(model2, test, ~ rmse(.x, .y)),
    model3_rmse = map2_dbl(model3, test, ~ rmse(.x, .y))
  ) %>% 
  select(model1_rmse, model2_rmse, model3_rmse) %>%
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse"
  ) %>%
  mutate(
    model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin(alpha = 0.6) +
  labs(
    title = "Cross-Validated RMSE for Three Models",
    x = "Model",
    y = "RMSE"
  )

cv_results
```

Based on these violin plots, my proposed model has a very high RMSE, meaning that there are much larger prediction errors and that it is not a good model for birthweight. Comparatively, the other two models have lower RMSEs, indicating that they are better predictive models.