---
title: "Homework 6"
author: "Mukta Patwari"
date: "2025-11-11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1234)
```

## Problem 1

Loading the homicide data

```{r}
library(readr)
library(broom)

homicide_data =
  read_csv("homicide-data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    solved = factor(
  case_when(
    disposition %in% c("Closed without arrest", "Open/No arrest") ~ "Unsolved",
    TRUE ~ "Resolved"),
  levels = c("Unsolved", "Resolved"))
  ) %>% 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
  ) %>% 
  mutate(
    victim_age = as.numeric(victim_age)
  )
```

Odds ratios for Baltimore

```{r}
baltimore_df =
  homicide_data %>% 
  filter(city_state == "Baltimore, MD")

baltimore_fit =
  baltimore_df %>% 
  glm(
    formula = solved ~ victim_age + victim_sex + victim_race,
    data = .,
    family = binomial
  ) 

baltimore_fit %>% 
  broom::tidy(conf.int = TRUE) %>% 
  mutate(
    OR = exp(estimate), 
    OR_conf_low = exp(conf.low), 
    OR_conf_high = exp(conf.high)) %>%
  filter(term == "victim_sexMale") %>%
  select(term, OR, OR_conf_low, OR_conf_high) %>% 
  knitr::kable(digits = 3)
```

Odds ratios for all cities

```{r}
cities_fit =
  homicide_data %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    fit = map(data, ~ glm(
      solved ~ victim_age + victim_sex + victim_race,
      data = .x,
      family = binomial
    )),
    
    tidy = map(fit, ~ 
      broom::tidy(.x, conf.int = TRUE) %>%
        mutate(
          OR = exp(estimate),
          OR_conf_low = exp(conf.low),
          OR_conf_high = exp(conf.high)
        ) %>%
        filter(term == "victim_sexMale") %>%
        select(term, OR, OR_conf_low, OR_conf_high)
    )
  ) %>%
  unnest(tidy)
```

Plot

```{r}
cities_fit %>%
  arrange(OR) %>%
  mutate(city_state = factor(city_state, levels = city_state)) %>%
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = OR_conf_low, ymax = OR_conf_high), width = 0.2) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(
    title = "Adjusted Odds Ratios for Solved Homicides",
    x = "City",
    y = "Adjusted Odds Ratio (Male vs Female)"
  )
```

## Problem 2

```{r}
library(p8105.datasets)
data("weather_df")
```

**Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution for your estimates, and describe these in words.**

```{r}
library(modelr)
library(patchwork)

bootstrap_weather =
  weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    model = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    r_squared = map_dbl(model, ~ glance(.x)$r.squared),
    beta_ratio = map_dbl(model, ~ {
      coefs = tidy(.x)
      b1 = coefs$estimate[coefs$term == "tmin"]
      b2 = coefs$estimate[coefs$term == "prcp"]
      b1 / b2
    })
  )

r_squared_plot =
  bootstrap_weather %>%
  ggplot(aes(x = r_squared)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Bootstrap Distribution of R-squared", 
    x = "R-squared", 
    y = "Density")

beta_ratio_plot =
  bootstrap_weather %>% 
  ggplot(aes(x = beta_ratio)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Bootstrap Distribution of Beta 1 / Beta 2",
    x = "Beta 1 / Beta 2",
    y = "Density"
  )

r_squared_plot + beta_ratio_plot
```

This plot shows

**Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval.**

```{r}
bootstrap_weather %>% 
  summarise(
    rsquared_lower = quantile(r_squared, 0.025),
    rsquared_upper = quantile(r_squared, 0.975),
    betaratio_lower = quantile(beta_ratio, 0.025),
    betaratio_upper = quantile(beta_ratio, 0.975)
  ) %>%
  pivot_longer(
    everything(),
    names_to = c("Estimates", "CI"),
    names_sep = "_"
  ) %>%
  pivot_wider(
    names_from = CI,
    values_from = value
  ) %>%
  knitr::kable(digits = 3)
```

## Problem 3

```{r}
birthweight_df =
  read_csv("birthweight.csv")
```

